---
title: "R - Python interoperability"
output:
  html_document:
    df_print: paged
---

This notebook shows how to combine R & Python via the `{reticulate}` package.\
This will require an installation of Python.

```{r, message = F, warning = F}
if(!require(reticulate)){install.packages("reticulate")}
library(tidyverse)
library(reticulate)
```

The reticulate package allows to manipulate objects in both languages:\
- if py_data is a Python object, you can access it in an R chunk via py\$py_data\
- if r_data is an R object, you can access it in a Python chunk via r.r_data

# Pure python first

To install/uninstall python modules, it is possible to proceed in R...

```{r}
#reticulate::conda_remove('r-reticulate', packages = c('numpy'))
#py_install("numpy")     # if you need to install python modules
#py_install("matplotlib")
#py_install("seaborn")
```

or in Python. Sometimes, you have to test both to make things work (module versioning conflicts, etc.).

```{python}
# Operates in /Users/coqueret/Library/r-miniconda-arm64/envs/r-reticulate/lib/python3.8/site-packages/
# !pip uninstall pandas -y
# !pip uninstall numpy -y
# !pip install numpy==1.24.2
# !pip install matplotlib
# !pip install pandas
# !pip install seaborn
# !pip uninstall six -y
# !pip install six==1.15.0
```

We fetch data from the Berkeley Earth organization (the NASA website is not 100% reliable strangely).

```{python, message = F, warning = F}
import pandas as pd                      # This library is for data wrangling
import numpy as np                       # Basic math library
import seaborn as sns                    # This is one library for plotting
import matplotlib.pyplot as plt          # Yet another library for visualization
#url = 'https://data.giss.nasa.gov/gistemp/graphs/graph_data/Global_Mean_Estimates_based_on_Land_and_Ocean_Data/graph.txt'
url = 'https://berkeley-earth-temperature.s3.us-west-1.amazonaws.com/Global/Land_and_Ocean_complete.txt'
data_temp = pd.read_csv(url, skiprows = 84, nrows = 2088, sep='\r\t', engine='python')  
data_temp = data_temp.iloc[0:,0].str.split(expand = True)
data_temp = data_temp.rename(columns = {0: 'Year', 1: 'Month', 2: 'Anomaly', 3: 'Uncertainty', 4: 'Ann_anomaly'})
data_temp = data_temp[['Year', 'Month', 'Anomaly', 'Uncertainty', 'Ann_anomaly']]         # Keep only 3 columns
data_temp = data_temp.query('Ann_anomaly != "NaN"') 
data_temp['Anomaly'] = pd.to_numeric(data_temp['Anomaly'])                                # Translate to number
data_temp['Uncertainty'] = pd.to_numeric(data_temp['Uncertainty'])                        # Translate to number
data_temp['Ann_anomaly'] = pd.to_numeric(data_temp['Ann_anomaly'])                        # Translate to number
data_temp['Date'] = data_temp['Year'].astype(str) + data_temp['Month'].astype(str) + '15' # Create date
data_temp['Date'] = pd.to_datetime(data_temp['Date'], format = '%Y%m%d')
data_temp
```

```{python}
plt.clf()
plt.plot(data_temp.Date, data_temp.Anomaly, label = "Temperature")
plt.plot(data_temp.Date, data_temp.Ann_anomaly, label = "Ann_anomaly")
plt.legend()
plt.show()
plt.rcParams["figure.figsize"]=(10, 6)
```

# Recycling python dataframes to use them in R

```{r}
df <- py$data_temp
df %>% ggplot(aes(x = Date, y = Ann_anomaly)) + geom_line() + theme_bw()
```

# Starting from R

Now we do the opposite: we start from R.

```{r}
library(gapminder)
data("gapminder")
force(gapminder)  # this step loads the data in the environment
```

Now let's switch to python! To access the R dataframe, we simply refer to "r.gapminder":

```{python}
pd.pivot_table(r.gapminder, values = 'pop', index = 'continent', aggfunc=np.sum).reset_index()
```

```{python}
pt = pd.pivot_table(r.gapminder, 
                    values = 'lifeExp', 
                    index = ['continent','year'], 
                    aggfunc=np.mean).reset_index()
pt
```

```{python}
#plt.clf()
sns.lineplot(data = pt, 
             x = "year", 
             y = "lifeExp",
             hue = 'continent')
#plt.show()
```

```{python}
import plotly.express as px
fig = px.line(pt, 
              x="year", 
              y="lifeExp", 
              color ='continent', 
              title='Life expectancy dynamics')
fig.show()
```

```{r}
pip install tensorflow
pip install keras

library(keras)
install_keras()

library(tensorflow)
install_tensorflow()
```
